An Elastica-driven Digital Curve Evolution Model for Image Segmentation\thanks{This work has been partially funded by CoMeDiC ANR-15-CE40-0006 research grant.}

### 2019-12-17 jmiv.tex
Simplification of opt vars definition
Mispelling corrections; bib file for reviewers answers

Geometric quantities are particularly useful as regularizers in
low-level image analysis, especially when little prior information is
known about the shape of interest. Length penalization is a
well-behaved, general purpose regularizer and many models in the
literature make use of it, from active contours
\cite{caseles97geodesic} to level-set formulations
\cite{malladi1995image,malladi1995shape}. Discrete graph-based
variational models have been particularly succesful to incorporate
length penalization as a penalizer, while keeping the ability to
extract a global optimum \cite{boykov01graphcut,appleton05geodesic}.

However length regularization shows limitations when segmenting small
or thin and elongated objects, as it tends to shrink solutions or
yields disconnected solutions. Such drawbacks can be problematic in
image segmentation, image restoration or image inpainting. It is thus
rather natural to consider curvature (especially squared curvature) as
a potential regularizer. Energies involving both length and squared
curvature are often called the {\em Elastica} model (that dates back
to Euler). The {\em Willmore energy} is its $n$-dimensional
extension. These were brought to attention in computer vision by
Mumford \cite{mumford1994elastica}. A similar concept is also present
in the second-order regularizer of the original snake model
\cite{kass1988snakes}.Indeed, an explicit curvature term was often
employed in early level-set methods,
like~\cite{malladi1995image,malladi1995shape,ballester01filljoint} for
segmentation or inpainting. However, \revision{Caselles} {\em et al.}
in~\cite{caseles97geodesic} observed that these terms as employed
were not geometric, in the sense that they depend on the
discretization parameters.

\subsection{Existing works}
The use of curvature for surface regularization in the case of the
Willmore energy was studied in~\cite{bobenko2005discrete}. Although
results look promising, authors used an inexact linearized version of
the square curvature at every step.

One of the first successful uses of geometric, squared curvature in
image processing is the inpainting algorithm described in
\cite{masnou98inpainting}, where authors evaluate the absolute
curvature along the level lines of a simply-connected image to
reconstruct its occluded parts. The non-intersection property of level
lines induces the construction of an efficient dynamic programming
algorithm. The curvature energy is nevertheless only coarsely
approximated. We may quote \cite{chan02elasticainpainting} as another
geometric inpainting method involving Elastica, which is optimized with
level-sets.

In image segmentation, Zehiry and Grady have shown that injecting
curvature as a regularizer can help recover thin and elongated objects
\cite{zehiry10fast}. Similarly Schoenemann {\em et al.}
\cite{schoenemann09linear} have considered the Elastica energy in
their ratio-based image segmentation
approach. In~\cite{schoenemann2011elastic}, Schoenemann~{\em et al.}
extend ratio-cut approaches to segmentation with a curvature
term. Their global optimization framework shows the considerable
advantages of using such regularizer in common binary image
segmentation tasks.  However the time complexity of their algorithm
is prohibitive, even if curvature is again coarsely
approximated.  In a more recent work, Lim {\em et al.} propose an
edge-weighted Willmore-like flow to segment
vertebra~\cite{lim2012introducing} using a level-set
formulation. They observe improved results compared
to~\cite{caseles97geodesic} and other perimeter-based priors, but
observe that a shape prior and a close-enough initialization must also be
provided to achieve good results.

In fact, it is still a very challenging task to efficiently handle
curvature in the context of image segmentation.  State-of-art methods
are in practice difficult to optimize and do not scale easily
\cite{zehiry10fast,schoenemann09linear,strandmark11globalframework,nieuwenhuis14efficient}. In
order to achieve reasonable running times, such approaches make use of
coarse curvature estimations for which the approximation error is
often unknown. Improving the quality of the curvature estimator has an
important impact on the accuracy of the results, but is
computationally too costly in these methods.

Segmentation methods using Elastica or Willmore energies can be
formulated via the computation of the corresponding flow,
i.e. following its gradient. Well-founded level-set methods involving
Willmore energy \cite{droske2004level} have been used in medical image
segmentation \cite{lim2012introducing}. Another level-set formulation
similar to the Chan-Vese model was proposed in
\cite{zhu2013image}. Willmore flow can also be carried out by
phase-field models (see \cite{bretin2015phase}) but they are less
suited to image segmentation since interfaces are blurred in such
models. Threshold dynamics can also be considered for such energies
\cite{esedoglu2008threshold} and has been proposed for image
\revision{disocclusion} \cite{esedoglu2005threshold}. However, flow-based methods
will always deliver a local minimum, which heavily depend on
initialization for good results. A similar approach is to formulate
the Elastica as a regularization in a variational framework and use
operator splitting to optimize it. Tai {\em et
  al.}~\cite{tai11elastica} and more recently Deng~{\em et
  al.}~\cite{deng2019new} have proposed this approach, which is
well-suited to inverse problem solving. The problem remains non-convex
and the nature of the converged iterates are not studied, i.e. how
close they are to a global optimum.

In contrast, researchers have sought convex relaxation of curvature-related
formulation. In~\cite{goldluecke11totalcurvature}, Goldluecke and Cremers introduce the {\em Total Curvature}, based
on the Menger-Melnikov curvature of a measure. The formulation is non-convex but can be relaxed into a well-correlated
but non-tight convex one. The relaxed formulation can be used to solve inverse problems including segmentation. However computing a solution is still expensive, requiring significant parallelisation efforts with a GPU to achieve
  acceptable computing times. In~\cite{bredies15convex}, Bredies {\em et al.} study a convex approximation of the
  Elastica energy, using a lifting scheme. Thanks to the lifting, it is not restricted to binary sets. It shows
  promising results and can be used for image restoration in addition to inpainting and segmentation. The scheme is
  however neither tight nor exact, and required 4 dimensions for a 2D approximation. In addition, it does not carry over
to $nD$. It also remains computationally expensive.

@@

%%\item{What are the difficulties in use curvature as a regularizer (point out convex relaxation approaches) and how a digital approach may help us }
%%\item{Outline: We present two models that illustrate how digital estimators can be used to minimize the elastica energy on digital sets.  }
%%\item{\todo{Add related works : geometric priors, binary segmentation, other digital curve evolution models}}
%%\item{\todo{Talk about Willmore flow with phase field or with threshold dynamics}}

@@

In this paper, we propose to investigate the use of the digital
integral invariant curvature estimator \cite{coeurjolly13integral} in
a discrete variational segmentation framework. More precisely, we show
how to incorporate it in a digital flow minimizing its squared
curvature. This model is expressible as a discrete combinatorial
optimization model, which we solve using a classical variant of QPBO
algorithm \cite{rother07qpbo}. Our solution does not compute a global
solution to the whole domain but a global solution to a band
surrounding a given shape. Our approach thus leads to a digital flow
similar to continuous Elastica, as shown by our experiments. We
present an application of this model as a post-processing step in a
segmentation framework, and demonstrate how it can improve standard
results. This paper is an extended version of a previous work that
appeared in \cite{antunes19}. We have included a deeper discussion on
our optimization method and we present a more thorough comparison of
our segmentation method with two other related works. Moreover, we
have included a new local combinatorial optimization approach for
elastica minimization, which show that our discrete model is
well-founded. Finally, \revision{the code related to} this paper is freely
available on github.\footnote{https://www.github.com/danoan/BTools}

\subsection{Outline}
Section~2 reviews the concept of multigrid convergence and highlights
its importance for the definition of digital estimators. We describe
two convergent estimators used in this paper, one that approaches the
length of elementary discrete contour elements, and the other that
approaches the curvature of a discrete contour. They are used in the
optimization model and in the definition of the digital
elastica. Section~3 describes a local combinatorial optimization model
suitable for several curvature estimators, that presents interesting
results but is too costly in practice. Section~4 describes the
proposed Elastica-driven curvature evolution model along with several
illustrations of digital flows. Section~5 explains how to use this
evolution model as a post-processing step in an image segmentation
framework and compares it to two related approaches. We conclude and
draw some perspectives to this work in Section~6.

@@

\subsection{Integral Invariant Curvature Estimator}
Generally speaking, an invariant is a function whose value is unaffected by the action
of some group on the elements of the domain.
%% Jaco: trop général, et pas utilisé après
%% Generally speaking, an invariant $\sigma$ is a real-valued function from some space $\Omega$ which value is unaffected by the action
%% of some group $\mathfrak{G}$ on the elements of the domain
%% \begin{align*}
%%   x \in \Omega, g \in \mathfrak{G}, \sigma(x) = v \longleftrightarrow \sigma(g \cdot x ) = v.
%% \end{align*}
Perimeter and curvature are examples of invariants for shapes on
$\mathbb{R}^2$ with respect to the euclidean group of rigid
transformations. Definition of integral area invariant and its link
with curvature is proven in \cite{manay04intinvariant}.

@@

In the remaining of this article, we are considering in the minimization of the Elastica energy. Given a Euclidean shape $X$ with smooth enough boundary, the Elastica is defined as
\begin{align}
  E(X) = \int_{\partial X}{(\alpha + \beta \kappa^2) ds}, \quad \text{for~} \alpha \ge 0, \beta \ge 0.
  \label{eq:elastica}
\end{align}

The digital version of Elastica energy approaching the continuous Elastica energy is defined as follows (it uses multigrid convergent estimators):
\begin{align}
	\hat{E}( D_h(X) ) = \sum_{\vec{e} \in \partial D_h(X)}{ \hat{s}(\vec{e})\left(\; \alpha + \beta \hat{\kappa}_{r}^2(D_h(X),\dot{\vec{e}},h) \; \right)},
	\label{eq:digital-energy}
\end{align}
where $\dot{\vec{e}}$ denotes the center of the edge $\vec{e}$. In the
expression above, we will substitute an arbitrary subset \revision{$S$} of
$\mathbb{Z}^2$ to $D_h(X)$ since the continuous shape $X$ is unknown.
In the following we omit the grid step $h$ to simplify expressions
(or, putting it differently, we assume that the shape of interest is
rescaled by $1/h$ and we set $h=1$).

@@

Given a digital shape $S^{(0)}$ we describe a process that generates a
sequence $S^{(i)}$ of shapes with non-increasing Elastica energy. The
idea is to define a neighborhood of shapes $\mathcal{N}^{(i)}$ to the
shape $S^{(i)}$ and choose the element of $\mathcal{N}^{(i)}$ with
lowest energy.  The process is suited for the integral invariant
estimator but also for other curvature estimators, for example, MDCA
\cite{roussillon11mdca}. As a matter of fact, our experiments have
shown that either estimators induce similar results.

@@

 \Input{A digital set $S$; the neighborhood parameter $n$; the maximum number of iterations \MIt; and a stop condition \Tol}
 \BlankLine
 \Delta $\longleftarrow$ \Tol+1\;
 $i \longleftarrow 0$\;
 \While{ \It $<$ \MIt \bf{and} \Delta $>$ \Tol  }{
  	\For{$ X \in \mathcal{N}_n(S^{(i)}) $}
	{
		\If{ $\hat{E}(X)$ $<$ $\hat{E}(X^\star)$ }
		{
			$X^\star \longleftarrow X$
		}
	}
	\It $\longleftarrow$ \It $+1$\;
	$S^{(i)} \longleftarrow X^\star$\;
	\Delta $\longleftarrow$ $\hat{E}(S^{(i-1)}) - \hat{E}(S^{(i)})$\;	
 }
 \caption{Local combinatorial optimization for elastica minimization.}
 \label{alg:local-search} 
\end{algorithm}

@@

The running time of algorithm \ref{alg:local-search} is summarized in table \ref{tab:summary-local-comb-rtime}. All the
experiments in this paper were executed on a \revision{$32$-core $2.4Ghz$ CPU and the number of pixels in the square and
flower shapes are respectively $841,1867$ for grid step $h=1.0$}. Although its use in practical applications is
limited, we demonstrate that digital estimators are effective in their measurements and the flows evolve as expected. \revision{In particular, the digital elastica energy for the experiments in figure \ref{fig:local-comb-square-results} approaches the global optimum (figure  \ref{fig:plot-elastica-local-search})}. We
observe that it is a complete digital approach, and we do not suffer from discretization and rounding problems, a common
issue in continuous models.  Furthermore we have checked that this approach works indifferently with Integral Invariant
curvature estimator and Maximal Digital Circular Arc curvature estimator. So the convergence of the digital curvature
estimator seems to be the cornerstone to get a digital curve behaving like a continuous Elastica.  In the next section,
we explore a more efficient approach.

\begin{figure}[!h]
\begin{minipage}[b]{0.5\textwidth}
\center
\includegraphics[scale=0.35]{images/local_search/square-flower.eps}
\caption{Elastica computation along iterations of algorithm \ref{alg:local-search}. \revision{The final shape has energy value closer to the optimal value $2\pi/5$ (dashed line)}.}
\label{fig:plot-elastica-local-search}
\end{minipage}\hspace{40pt}%
\begin{minipage}[b]{0.45\textwidth}
\revision{
\center
\captionsetup{type=table}
\begin{tabular}{|l|c|c|}
\hline
& $h=1.0$ & $h=0.5$\\
\hline
Square & 4s (0.4s/it) & 77s (3s/it)\\
\hline
Flower & 252s (4s/it)  & 3246s (42s/it)\\
\hline
\end{tabular}
}
\caption{Running times for the local combinatorial optimization algorithm with $\revision{ 0 \leq n \leq 50}$. Thirty two threads were used.}
\label{tab:summary-local-comb-rtime}
\end{minipage}
\end{figure}

@@

%	Let $\Omega_C$ be the collection of connected digital sets in $\Omega$. Given a digital shape $S$, we wish to evolve it into a shape $S^\star$ similar to $S$ but with lower digital elastica energy. We recall the digital curvature estimator definition
%	
%\begin{align}
%	\hat{\kappa}^2(y) &= c_1\Big( c_2 - | B_r(y) \cap S | \Big)^2, 
%	\label{eq:curvature-estimator-pixels}
%\end{align}
%
%where $c_1=3/r^6$ and $c_2=\pi r^2/2$. Let $X,Y$ be the set of pixels and linels, respectively, in $\Omega$. We define the following binary model
%		
%\begin{align}
%\text{Minimize}_{x,y} & \sum_{y}{c_1y \Big(c_2 - \sum_{x_j \in B_r(y)}{x_j}\Big)^2	 } + F(S,X_{=1} \cup Y_{=1}) \nonumber \\
%\text{subject to} & \nonumber \\
%& (X_{=1} \cup Y_{=1}) \in \Omega_C, \; x,y \in \{0,1\}^\Omega
%\label{eq:general-binary-formulation}
%\end{align}
%
%
%where $F$ is some measure of similarity, e.g. distance, between initial shape $S$ and candidate solution. Such term is needed, otherwise the uninteresting empty set would be the optimal solution.
%
%Formulation \eqref{eq:general-binary-formulation} is of order four, binary and non-convex. These properties make it a hard optimization problem. We are going to explore an alternative energy, which is a simplification of the former.

%\subsection{A first model}
%
%We note that the elastica can be decreased by attenuating non-flat regions. The idea is to evolve a flow $S^{(i)}$ where $S^{(i+1)}$ is derived from the minimization of the sum of the squared curvature at $S^{(i)}$. We exchange solving a fourth order energy to solve multiple second order problems. We still have to guarantee that solutions are connected. We handle this issue by limiting the optimization region to a subset of $\Omega$, namely the inner pixel boundary of  $S^{(i)}$.

@@

\revision{We observe that the integral invariant estimator \eqref{eq:curvature_approximation} is originally designed to be evaluated on the digital boundary of the shape, which is the unknown of our problem. Injecting contour information leads to a third order model, which we prefer to avoid. Instead, we are going to exploit the II estimator definition in order to compute a flow that evolves proportionally to the squared curvature of the current shape.}

@@

\revision{We assume an ordering in $\Omega$, i.e., there exists a bijective function $\omega : \Omega \rightarrow \{1 \cdots |\Omega| \}$. Moreover, we associate to every pixel in $\Omega$ a binary variable grouped in the set $X(\Omega)$}

\begin{align*}
	\revision{X(\Omega) := \left\{ x_{\omega(p)} \in \{0,1\} \; | \; p \in \Omega \right\}.}
\end{align*}

\revision{We extend the definition for every subset $P \in \Omega$, i.e.,}

@@

%%\jaco{Motivation in digital geometry of why it makes sense to compute things farther.}

@@

{In our experiments, the best results are obtained by executing DCE algorithm with $m$ equal to $r$, where $r$ is
  the estimation ball radius (see figure \ref{fig:mx-square-flow}). We observe that Elastica may increase after some
  iterations if chosen radius is too large, as in the case of the triangle in figure \ref{fig:mx-flow-gs-radius-effect}
  in which the flow converges to a single point. We conjecture that an appropriated value for the radius should be given
  by the shape reach.  The produced flow has no difficulties in handling changes on topology, and it presents different
  speeds for regions with low and high curvature values, as illustrated in figure
  \ref{fig:mx-speed-variation-hole-filling}.

@@

		\setcounter{subfigure}{-2}
		\subfloat[$r=5,h=0.5$]{%
		\subfloat{%
		\includegraphics[scale=0.2]{images/flow/grid-radius-effect/triangle/r5-h0.5/summary_flow.eps}
		}%1
		\hspace{15pt}
		\subfloat{%
		\includegraphics[scale=0.15]{images/flow/grid-radius-effect/square/r5-h0.5/summary_flow.eps}}%
		\hspace{15pt}
		\subfloat{%
\includegraphics[scale=0.2]{images/flow/grid-radius-effect/flower/r5-h0.5/summary_flow.eps}}%
		}\\
		\subfloat[\revision{Elastica evaluation}]{%
		\includegraphics[scale=0.5]{images/flow/elastica-energy-plot/square-flower-triangle.eps}						
		}
\caption{The choice of radius impacts the flow. In the figures, the flow ceases to evolve for all shapes  when $r=3$ (a). In figure (b), for $r=5$, the triangle evolves to a single point, while the others stop  in an intermediate shape, as in (a). In figure (c), we observe that for a given choice of radius, the elastica may increase after a certain number of iterations. }
\label{fig:mx-flow-gs-radius-effect}
\end{figure}

@@

We present an application of our digital curve evolution algorithm to supervised image segmentation. The DCE acts as a
contour correction method. Here we use a data fidelity term in order to characterize the object of
interest. Given foreground and background seeds selected by the user, we derive mixed Gaussian distributions  of color
intensities $G_f$ and $G_b$, and we define the data fidelity term as the cross-entropy, i.e.

@@

We have studied in depth several digital curve evolution models based on a digital version of the Elastica energy and we
have presented an application to image segmentation. The processes we have described are completely digital, and do not
suffer from issues that typically arise in models that passes through a discretization stage, such as
rounding. Moreover, the model can handle changes in topology and its results are competitive with similar approaches
while achieving reasonable running times.

@@

%% for reference

%% In this work, we investigate the use of a more suitable curvature
%% estimator with multigrid convergent property and its application as a
%% boundary regularizer in a digital flow minimizing its squared
%% curvature. Our method decreases the elastica energy of the contour and
%% its evolution is evaluated on several digital flows.

%% Finally, we
%% present an application of the model as a post-processing step in a
%% segmentation framework. The code is freely available on
%% github\footnote{https://www.github.com/danoan/BTools}.

%\begin{itemize}
%% \item{Geometric properties and regularization}
%% \item{The role of curvature}

%  Quote \cite{schoenemann2011elastic} for the interest of using curvature energy when segmenting regions in gradient-based data fitting term. It can look for global optimum within a graph representation of the image. Main problem of their approach is the computational complexity as well as rather coarse approximations of curvature. (Be more precise here).

%%\item{\textbf{Classical work on elastica and Wilmore flow in imaging}}
  %%\begin{itemize}
  %% \item{Level lines based disocclusion \cite{masnou98inpainting} }	%% DONE
 %%  \item{Filling-In by Joint Interpolation of Vector Fields and Gray Levels \cite{ballester01filljoint} } %% DONE
 %%  \item{Eulers Elastica and Curvature Based Inpaintings \cite{chan02elasticainpainting} } %% DONE
 %% \item{Discrete willmore flow \cite{bobenko2005discrete}}
 %%  \item{Introducing Willmore flow into level set segmentation of spinal vertebrae \cite{lim2012introducing}}	 %%% DONE
 %% \end{itemize}

%%\item{\textbf{Level-sets}}
 %% \begin{itemize}
  %% \item{Image processing via level set curvature flow \cite{malladi1995image} } %% DONE
  %% \item{Shape Modeling with Front Propagation: A Level Set Approach \cite{malladi1995shape} }	%% DONE
%%	\end{itemize}

%%\item{\textbf{Convex relaxation}}
%%  \begin{itemize}
%%  \item{A convex lower semicontinuous approximation of Eulers elastica energy \cite{bredies15convex}}
 %% \item{Introducing Total Curvature for Image Processing \cite{goldluecke11totalcurvature}}
 %% \end{itemize}

>>>>

Geometric quantities are particularly useful as regularizers in
low-level image analysis, especially when little prior information is
known about the shape of interest. Length penalization is a
well-behaved, general\revision{-}purpose regularizer and many models in the
literature make use of it, from active contours
\cite{caseles97geodesic} to level-set formulations
\cite{malladi1995image,malladi1995shape}. Discrete graph-based
variational models have been particularly succes\revision{s}ful to incorporate
length penalization as a penalizer while keeping the ability to
extract a global optimum \cite{boykov01graphcut,appleton05geodesic}.

However, length regularization shows limitations when segmenting small
or thin and elongated objects, as it tends to shrink solutions or
yields disconnected solutions. Such drawbacks can be problematic in
image segmentation, image restoration or image inpainting. It is thus
rather natural to consider curvature (especially squared curvature) as
a potential regularizer. Energies involving both length and squared
curvature are often called the {\em Elastica} model (that dates back
to Euler). The {\em Willmore energy} is its $n$-dimensional
extension. These were brought to attention in computer vision by
Mumford \cite{mumford1994elastica}. A similar concept is also present
in the second-order regularizer of the original snake model
\cite{kass1988snakes}. Indeed, an explicit curvature term was often
employed in early level-set methods,
like~\cite{malladi1995image,malladi1995shape,ballester01filljoint} for
segmentation or inpainting. However, \revision{Caselles} {\em et al.}
in~\cite{caseles97geodesic} observed that these terms as employed
were not geometric, in the sense that they depend on the
discretization parameters.

\subsection{Existing works}
The use of curvature for surface regularization in the case of the
Willmore energy was studied in~\cite{bobenko2005discrete}. Although
results look promising, \revision{the} authors used an inexact linearized version of
the square curvature at every step.

One of the first successful uses of curvature in
image processing is the inpainting algorithm described in
\cite{masnou98inpainting}, where authors evaluate the absolute
curvature along the level lines of a simply-connected image to
reconstruct its occluded parts. The non-intersection property of level
lines induces the construction of an efficient dynamic programming
algorithm. The curvature energy is nevertheless only coarsely
approximated. We may quote \cite{chan02elasticainpainting} as another
geometric inpainting method involving Elastica, which is optimized with
level-sets.

In image segmentation, Zehiry and Grady have shown that injecting
curvature as a regularizer can help recover thin and elongated objects
\cite{zehiry10fast}. Similarly, Schoenemann {\em et al.}
\cite{schoenemann09linear} have considered the Elastica energy in
their ratio-based image segmentation
approach. In~\cite{schoenemann2011elastic}, Schoenemann~{\em et al.}
extend ratio-cut approaches to segmentation with a curvature
term. Their global optimization framework shows the considerable
advantages of using such regularizer in common binary image
segmentation tasks.  However, the time complexity of their algorithm
is prohibitive, even if curvature is again coarsely
approximated.  In a more recent work, Lim {\em et al.} propose an
edge-weighted Willmore-like flow to segment
vertebra~\cite{lim2012introducing} using a level-set
formulation. They observe improved results compared
to~\cite{caseles97geodesic} and other perimeter-based priors, but
observe that a shape prior and a close-enough initialization must also be
provided to achieve good results.

In fact, it is still a very challenging task to efficiently handle
curvature in the context of image segmentation.  State-of-art methods
are in practice difficult to optimize and do not scale easily
\cite{zehiry10fast,schoenemann09linear,strandmark11globalframework,nieuwenhuis14efficient}. In
order to achieve reasonable running times, such approaches make use of
coarse curvature estimations for which the approximation error is
often unknown. Improving the quality of the curvature estimator has an
important impact on the accuracy of the results, but is
computationally too costly in these methods. \revision{Moreover, parametric approaches to Elastica  as snakes and splines depend on the choice of anchor points and  multigrid convergence hasn't been proven for them.}

Segmentation methods using Elastica or Willmore energies can be
formulated via the computation of the corresponding flow,
i.e. following its gradient. Well-founded level-set methods involving
Willmore energy \cite{droske2004level} have been used in medical image
segmentation \cite{lim2012introducing}. Another level-set formulation
similar to the Chan-Vese model was proposed in
\cite{zhu2013image}. Willmore flow can also be carried out by
phase-field models (see \cite{bretin2015phase}) but they are less
suited to image segmentation since interfaces are blurred in such
models. Threshold dynamics can also be considered for such energies
\cite{esedoglu2008threshold} and has been proposed for image
\revision{disocclusion} \cite{esedoglu2005threshold}. However, flow-based methods
will always deliver a local minimum, which heavily depend on
initialization for good results. A similar approach is to formulate
the Elastica as a regularization in a variational framework and use
operator splitting to optimize it. Tai {\em et
  al.}~\cite{tai11elastica} and more recently Deng~{\em et
  al.}~\cite{deng2019new} have proposed this approach, which is
well-suited to inverse problem solving. The problem remains non-convex
and the nature of the converged iterates is not studied, i.e. how
close they are to a global optimum.

In contrast, researchers have sought convex relaxation of curvature-related
formulation. In~\cite{goldluecke11totalcurvature}, Goldluecke and Cremers introduce the {\em Total Curvature}, based
on the Menger-Melnikov curvature of a measure. The formulation is non-convex but can be relaxed into a well-correlated
but non-tight convex one. The relaxed formulation can be used to solve inverse problems including segmentation. However computing a solution is still expensive, requiring significant parallelisation efforts with a GPU to achieve
  acceptable computing times. In~\cite{bredies15convex}, Bredies {\em et al.} study a convex approximation of the
  Elastica energy, using a lifting scheme. Thanks to the lifting, it is not restricted to binary sets. It shows
  promising results and can be used for image restoration in addition to inpainting and segmentation. The scheme is,
  however, neither tight nor exact, and required 4 dimensions for a 2D approximation. In addition, it does not carry over
to $nD$. It also remains computationally expensive.

@@

@@

In this paper, we propose to investigate the use of the digital
integral invariant curvature estimator \cite{coeurjolly13integral} in
a discrete variational segmentation framework. More precisely, we show
how to incorporate it in a digital flow minimizing its squared
curvature. This model is expressible as a discrete combinatorial
optimization model, which we solve using a classical variant of QPBO
algorithm \cite{rother07qpbo}. Our solution does not compute a global
solution to the whole domain but a global solution to a band
surrounding a given shape. Our approach thus leads to a digital flow
similar to continuous Elastica, as shown by our experiments. We
present an application of this model as a post-processing step in a
segmentation framework and demonstrate how it can improve standard
results. This paper is an extended version of a previous work that
appeared in \cite{antunes19}. We have included a deeper discussion on
our optimization method and we present a more thorough comparison of
our segmentation method with two other related works. Moreover, we
have included a new local combinatorial optimization approach for
Elastica minimization, which shows that our discrete model is
well-founded. Finally, \revision{the code related to} this paper is freely
available on GitHub.\footnote{https://www.github.com/danoan/BTools}

\subsection{Outline}
Section~2 reviews the concept of multigrid convergence and highlights
its importance for the definition of digital estimators. We describe
two convergent estimators used in this paper, one that approaches the
length of elementary discrete contour elements, and the other that
approaches the curvature of a discrete contour. They are used in the
optimization model and in the definition of the digital
Elastica. Section~3 describes a local combinatorial optimization model
suitable for several curvature estimators, that presents interesting
results but is too costly in practice. Section~4 describes the
proposed Elastica-driven curvature evolution model along with several
illustrations of digital flows. Section~5 explains how to use this
evolution model as a post-processing step in an image segmentation
framework and compares it to two related approaches. We conclude and
draw some perspectives to this work in Section~6.

@@

\subsection{Integral Invariant Curvature Estimator}
Generally speaking, an invariant is a function whose value is unaffected by the action
of some group on the elements of the domain.

Perimeter and curvature are examples of invariants for shapes on
$\mathbb{R}^2$ with respect to the Euclidean group of rigid
transformations. Definition of integral area invariant and its link
with curvature is proven in \cite{manay04intinvariant}.

@@

In the remaining of this article, we are considering the minimization of the Elastica energy. Given a Euclidean shape $X$ with smooth enough boundary, the Elastica is defined as
\begin{align}
  E(X) = \int_{\partial X}{(\alpha + \beta \kappa^2) ds}, \quad \text{for~} \alpha \ge 0, \beta \ge 0.
  \label{eq:elastica}
\end{align}

The digital version of Elastica energy approaching the continuous Elastica energy is defined as follows (it uses multigrid convergent estimators):
\begin{align}
	\hat{E}( D_h(X) ) = \sum_{\vec{e} \in \partial D_h(X)}{ \hat{s}(\vec{e})\left(\; \alpha + \beta \hat{\kappa}_{r}^2(D_h(X),\dot{\vec{e}},h) \; \right)},
	\label{eq:digital-energy}
\end{align}
where $\dot{\vec{e}}$ denotes the center of the edge $\vec{e}$. In the
expression above, we will substitute an arbitrary subset \revision{$S$} of
$\mathbb{Z}^2$ to $D_h(X)$ since the continuous shape $X$ is unknown.
In the following, we omit the grid step $h$ to simplify expressions
(or, putting it differently, we assume that the shape of interest is
rescaled by $1/h$ and we set $h=1$).

@@

Given a digital shape $S^{(0)}$ we describe a process that generates a
sequence $S^{(i)}$ of shapes with non-increasing Elastica energy. The
idea is to define a neighborhood of shapes $\mathcal{N}^{(i)}$ to the
shape $S^{(i)}$ and choose the element of $\mathcal{N}^{(i)}$ with
lowest energy.  The process is suited for the integral invariant
estimator but also for other curvature estimators, for example, MDCA
\cite{roussillon11mdca}. As a matter of fact, our experiments have
shown that either estimator induce similar results.

@@

 \Input{A digital set $S$; the neighborhood parameter $n$; the maximum number of iterations \MIt; and a stop condition \Tol}
 \BlankLine
 \Delta $\longleftarrow$ \Tol+1\;
 $i \longleftarrow 0$\;
 \While{ \It $<$ \MIt \bf{and} \Delta $>$ \Tol  }{
  	\For{$ X \in \mathcal{N}_n(S^{(i)}) $}
	{
		\If{ $\hat{E}(X)$ $<$ $\hat{E}(X^\star)$ }
		{
			$X^\star \longleftarrow X$
		}
	}
	\It $\longleftarrow$ \It $+1$\;
	$S^{(i)} \longleftarrow X^\star$\;
	\Delta $\longleftarrow$ $\hat{E}(S^{(i-1)}) - \hat{E}(S^{(i)})$\;	
 }
 \caption{Local combinatorial optimization for \revision{E}lastica minimization.}
 \label{alg:local-search} 
\end{algorithm}

@@

The running time of algorithm \ref{alg:local-search} is summarized in table \ref{tab:summary-local-comb-rtime}. All the
experiments in this paper were executed on a \revision{$32$-core $2.4Ghz$ CPU and the number of pixels in the square and
flower shapes are respectively $841,1867$ for grid step $h=1.0$}. Although its use in practical applications is
limited, we demonstrate that digital estimators are effective in their measurements and the flows evolve as expected. \revision{In particular, the digital elastica energy for the experiments in figure \ref{fig:local-comb-square-results} approaches the global optimum (figure  \ref{fig:plot-elastica-local-search})}. We
observe that it is a complete digital approach, and we do not suffer from discretization and rounding problems, a common
issue in continuous models.  Furthermore, we have checked that this approach works indifferently with Integral Invariant
curvature estimator and Maximal Digital Circular Arc curvature estimator. So the convergence of the digital curvature
estimator seems to be the cornerstone to get a digital curve behaving like a continuous Elastica.  In the next section,
we explore a more efficient approach.

\begin{figure}[!h]
\begin{minipage}[b]{0.5\textwidth}
\center
\includegraphics[scale=0.35]{images/local_search/square-flower.eps}
\caption{Digital elastica computation along iterations of algorithm \ref{alg:local-search}. \revision{The final shape has energy value closer to the optimal value $2\pi/5$ (dashed line)}.}
\label{fig:plot-elastica-local-search}
\end{minipage}\hspace{40pt}%
\begin{minipage}[b]{0.45\textwidth}
\revision{
\center
\captionsetup{type=table}
\begin{tabular}{|l|c|c|}
\hline
& $h=1.0$ & $h=0.5$\\
\hline
Square & 4s (0.4s/it) & 77s (3s/it)\\
\hline
Flower & 252s (4s/it)  & 3246s (42s/it)\\
\hline
\end{tabular}
}
\caption{Running times for the local combinatorial optimization algorithm with $\revision{ 0 \leq n \leq 50}$. Thirty two threads were used.}
\label{tab:summary-local-comb-rtime}
\end{minipage}
\end{figure}

@@

@@

\revision{We observe that the integral invariant estimator \eqref{eq:curvature_approximation} is originally designed to be evaluated on the digital boundary of the shape, which is the unknown of our problem. Injecting contour information leads to a third-order model, which we prefer to avoid. Instead, we are going to exploit the II estimator definition to compute a flow that evolves proportionally to the squared curvature of the current shape.}

@@

\revision{We assume an ordering in $\Omega$, i.e., there exists a bijective function $\omega : \Omega \rightarrow \{1 \cdots |\Omega| \}$. Moreover, we associate to any subset $P \subset \Omega$ the set of binary variables $X(P)$ defined as}

@@

@@

{In our experiments, the best results are obtained by executing DCE algorithm with $m$ equal to $r$, where $r$ is
  the estimation ball radius (see figure \ref{fig:mx-square-flow}). We observe that digital elastica may increase after some
  iterations if chosen radius is too large, as in the case of the triangle in figure \ref{fig:mx-flow-gs-radius-effect}
  in which the flow converges to a single point. We conjecture that an appropriated value for the radius should be given
  by the shape reach.  The produced flow has no difficulties in handling changes on topology, and it presents different
  speeds for regions with low and high curvature values, as illustrated in figure
  \ref{fig:mx-speed-variation-hole-filling}.

@@

		\setcounter{subfigure}{-2}
		\subfloat[$r=5,h=0.5$]{%
		\subfloat{%
		\includegraphics[scale=0.2]{images/flow/grid-radius-effect/triangle/r5-h0.5/summary_flow.eps}
		}%1
		\hspace{15pt}
		\subfloat{%
		\includegraphics[scale=0.15]{images/flow/grid-radius-effect/square/r5-h0.5/summary_flow.eps}}%
		\hspace{15pt}
		\subfloat{%
\includegraphics[scale=0.2]{images/flow/grid-radius-effect/flower/r5-h0.5/summary_flow.eps}}%
		}\\
		\subfloat[\revision{Digital elastica evaluation}]{%
		\includegraphics[scale=0.5]{images/flow/elastica-energy-plot/square-flower-triangle.eps}						
		}
\caption{The choice of radius impacts the flow. In the figures, the flow ceases to evolve for all shapes  when $r=3$ (a). In figure (b), for $r=5$, the triangle evolves to a single point, while the others stop  in an intermediate shape, as in (a). In figure (c), we observe that for a given choice of radius, the digital elastica may increase after a certain number of iterations. }
\label{fig:mx-flow-gs-radius-effect}
\end{figure}

@@

We present an application of our digital curve evolution algorithm to supervised image segmentation. The DCE acts as a
contour correction method. Here we use a data fidelity term in order to characterize the object of
interest. Given foreground and background seeds selected by the user, we derive mixed Gaussian distributions of color
intensities $G_f$,$G_b$, and we define the data fidelity term as the cross-entropy, i.e.

@@

We have studied in depth several digital curve evolution models based on a digital version of the Elastica energy and we
have presented an application to image segmentation. The processes we have described are completely digital and do not
suffer from issues that typically arise in models that pass through a discretization stage, such as
rounding. Moreover, the model can handle changes in topology and its results are competitive with similar approaches
while achieving reasonable running times.

@@
